---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
deaths <- av %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(Time = parse_number(Time))

head(deaths)
```

Similarly, deal with the returns of characters.

```{r}
returns <- av %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(Time = parse_number(Time))

head(returns)
```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
avg_deaths <- deaths %>%
  filter(Death == "YES") %>%
  group_by(URL) %>%
  summarise(total_deaths = n()) %>%
  summarise(avg = mean(total_deaths))

avg_deaths
```


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.


### Include the code

Make sure to include the code to derive the (numeric) fact for the statement


### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.


## Individually - Mariana Correa

### FiveThirtyEight Statement

> "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That's about 40 percent of all people who have ever signed on to the team."

### Include the code

```{r}
num_avengers <- nrow(av)
num_avengers

died_at_least_once <- deaths %>%
  filter(Death == "YES") %>%
  distinct(URL) %>%
  nrow()

died_at_least_once

pct_died <- died_at_least_once / num_avengers * 100
pct_died
```

### Include your answer

My calculations show that the dataset contains 173 Avengers, and 69 of them dies at least once, totaling about 40%. Therefore, the original statement is correct.


## Ewan Individual work

### Ewan FiveThirtyEight Statement

"There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife

## Ewan's work
```{r}
first_deaths <- deaths %>%
  filter(Death == "YES") %>%
  group_by(URL) %>%
  arrange(Time) %>%
  slice(1) %>%
  ungroup()

first_returns <- first_deaths %>%
   left_join(returns, by = c("URL", "Time")) %>%
  summarise(
  return_rate = mean(Return == "YES", na.rm = TRUE)
  )
first_returns
  
  
```

## Ewan's Answer
after fact checking the statement I found that the return rate of avengers after their first death is 66.67% which is a 2 in 3 chance, so the statement was correct.




### Mason's FiveThirtyEight Statement

> "Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics. In fact, Hawkeye died twice!"


### Mason's code

```{r}
deaths_filtered <- deaths %>% filter (Death == 'YES', Name.Alias %in% c('Anthony Edward "Tony" Stark','Robert Bruce Banner' ,'Steven Rogers', 'Thor Odinson', 'Clinton Francis Barton', 'Natalia Alianovna Romanova', 'Wanda Maximoff', 'Pietro Maximoff', 'Victor Shade (alias)'))
deaths_filtered

counts_df <- deaths_filtered |> count(Name.Alias)
counts_df
```


### Mason's answer

The quote stated that all 9 of the avengers have died at least once, with hawkeye having suffered two deaths. I filtered the deaths dataframe to only those 9 avengers and the rows in which they have a death marked, and summed them up in a dataframe which counts the deaths. The statement is TRUE! However, it fails to account for the fact that Thor has also died twice.

## Ryan's Individual work

### Ryan's FiveThirtyEight Statement

> "Given the Avengers' 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so."

### Ryan's code

```{r}
death_frequency <- deaths %>%
  filter(Death == "YES") %>%
  summarise(
    total_deaths = n(),
    years_operating = 53,
    months_per_death = (years_operating * 12) / total_deaths
  )

death_frequency
```

### Ryan's Answer

The data shows 89 total deaths across the 53 years the Avengers have been in operation. That works out to roughly one death every 7.15 months, which rounds to "every seven months or so." The FiveThirtyEight claim is correct.
